\label{section:outputs}
\setcounter{footnote}{0}
\section{Outputs}

For each inout file, \etwomsa\, produces the
following output files:

\begin{sreitems}{}

\item[\emprog{file.e2msa.AFG.sto}]
%
The resulting alignment in Stockholm format.

\item[\emprog{file.AFG.bench}] 
%
If the input file is an alignment, it also producds file and option \emprog{file.AFG.bench}. This file
provides information about the comparison of the input alignment with the \etwomsa\ alignment

\begin{sreoutput}
#MSA_NAME  TP_hom TRUE_hom FOUND_hom      
globins4.1 760 827 804 0 0 29.06 29.06 88.11 83.68 0.48 7.36 e2msa.AFG 171 174 147.25 5.06 147.25 5.06 6.0 1.41 4.5 1.91 3.96 2.71 5.94 3.04 -2288.18 0 0 0
\end{sreoutput}

 This file produces a one line output per alignment in the file.
 \begin{sreitems}{\prog{Second column}}
 \item[\prog{Column 1}]  MSA name.
 \item[\prog{Column 2}]  Number of true homologies found in the \etwomsa\ alignment (TP=true positives).
 \item[\prog{Column 3}]  Number of homologies in the input alignment (T=true homologies).
 \item[\prog{Column 4}]  Number of true homologies found in the \etwomsa\ alignment (F=found).
 \item[\prog{Column 5}]  Number of nonhomologous  positions in \etwomsa\ alignment.
 \item[\prog{Column 6}]  Number of nonhomologous  positions in input alignment.
 \item[\prog{Column 7}]  Average \% identity in input alignment.
 \item[\prog{Column 8}]  Average \% identity in \etwomsa\ alignment.
 \item[\prog{Column 9}]  Average number of matches in input alignment.
 \item[\prog{Column 10}] Average number of matches in \etwomsa\ alignment.
 \item[\prog{Column 11}] Average time of inferred tree.
 \item[\prog{Column 12}] Time used to run the \etwomsa\ program.
 \item[\prog{Column 13}] Evolutionary model used.
 \item[\prog{Column 14}] Length of input alignment.
 \item[\prog{Column 15}] Length of \etwomsa\ alignment.
 \item[\prog{Column 16-17}] Avg/standard-deviation of length of seqs in input alignment.
 \item[\prog{Column 18-19}] Avg/standard-deviation of length of seqs in \etwomsa\ alignment.
 \item[\prog{Column 20-21}] Avg/standard-deviation of number of gaps in input alignment.
 \item[\prog{Column 22-23}] Avg/standard-deviation of number of gaps in \etwomsa\ alignment.
 \item[\prog{Column 24-25}] Avg/standard-deviation of length of gaps in input alignment.
 \item[\prog{Column 26-27}] Avg/standard-deviation of length of gaps in \etwomsa\ alignment.
 \item[\prog{Column 28}]    Aignment score.
 \item[\prog{Column 29}]    Number of homologous  positions in input alignment.
 \item[\prog{Column 30}]    Number of homologous  positions in \etwomsa\ alignment.
 \item[\prog{Column 31}]    number of homologous  positions both in input alignment and \etwomsa\ alignment.
 \end{sreitems}

\end{sreitems}
