
\section{Some other topics}
\label{section:more}
\setcounter{footnote}{0}

\subsection{How do I cite \rscape?}

Rivas, E. \textit{et al.}, \textit{``A statistical test for conserved
  RNA structure shows lack of evidence for structure in lncRNAs''},
Nature Methods 14, 45–48 (2017).\\
\url{https://www.nature.com/articles/nmeth.4066}\\

\noindent
Rivas, E. and Eddy, S. E., \textit{``Response to Tavares et al.,
  “Covariation analysis with improved parameters reveals conservation
  in lncRNA structures”''}, (2018).\\
\url{https://doi.org/10.1101/2020.02.18.955047}.\\

\noindent
Rivas, E. \textit{et al.}, \textit{``Estimating the power of sequence
  covariation analysis for detecting conserved RNA structure''},
\textit{Bionformatics}, 36, 3072–3076, (2020).\\
\url{https://doi.org/10.1093/bioinformatics/btaa080}\\

\noindent
Rivas, E., \textit{``RNA covariation at helix-level resolution for the identification of evolutionarily conserved RNA structure''},
\textit{PLOS Comput Biol, 19(7): e1011262, (2023).}\\
  \url{https://doi.org/10.1371/journal.pcbi.1011262}.\\

\subsection{How do I cite CaCoFold?}

Rivas, E., \textit{``RNA structure prediction using
  positive and negative evolutionary information''},
\textit{PLOS Comput Biol}, 16(10), e1008387, (2020).\\
\url{https://doi.org/10.1371/journal.pcbi.1008387}.\\

\subsection{How do I cite CaCoFold-R3D?}

Karan, A and Rivas, E., \textit{``All-at-once RNA folding with 3D motif prediction framed by evolutionary information''},
\textit{bioRxiv}, Dec (2024).\\
\url{https://doi.org/10.1101/2024.12.17.628809}, 20 Dec 2024.\\


You should also cite what version of the software you used. We archive
all old versions, so anyone should be able to obtain the version you
used, when exact reproducibility of an analysis is an issue.

The version number is in the header of most output files. To see it
quickly, do something like \prog{R-scape -h} to get a help page, and
the header will say:

\begin{sreoutput}
# R-scape :: RNA Structural Covariation Above Phylogenetic Expectation
# R-scape 2.5.7 (March 2025)
# Copyright (C) 2017-2025 Elena Rivas, Harvard University.
# Freely distributed under the GNU General Public License (GPLv3).
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\end{sreoutput}

So (from the second line there) this is from \rscape\ v2.5.7.

\subsection{How do I report a bug?}

Email us, at \url{elenarivas@fas.harvard.edu}.

Before we can see what needs fixing, we almost always need to
reproduce a bug on one of our machines. This means we want to have a
small, reproducible test case that shows us the failure you're seeing.
So if you're reporting a bug, please send us:

\begin{itemize}
 \item A brief description of what went wrong.
 \item The command line(s) that reproduce the problem.
 \item Copies of any files we need to run those command lines.
 \item Information about what kind of hardware you're on, what
   operating system, and what compiler and version you used, with what
   configuration arguments.
\end{itemize}



