
\section{Tutorial for running R-scape}
\label{section:tutorial}
\setcounter{footnote}{0}

Here's a tutorial walk-through of how to use \rscape. This should
suffice to get you started.

\subsection {Modes of \rscape}

For an input alignment, \rscape\ reports all pairs that have
covariation scores with E-values smaller than a target E-value.\\

\noindent
R-scape has two different \textbf{modes} of operation which determine
how it calculates E-values, for which it needs to know how many
possible base pairs were tested (i.e. E-values are
multiple-test-corrected). The E-values are calculated in one of two ways:

\begin{tabular}{ll}
\multicolumn{2}{l}{\textbf{A one-set statistical test:} \textit{default}} \\ 
 & \\ 
\textbf{}   & E-values are calculated assuming that all pairs are possible.\\
\textbf{}   & This is the default behavior of \rscape.\\
 & \\ 
\multicolumn{2}{l}{\textbf{A two-set statistical test: } \prog{option -s}} \\ 
 & \\ 
\textbf{}   & If the alignment has associated a \emph{given structure}, \textbf{\prog{option -s}} performs two independent statistical tests: \\
\textbf{}   & one for the pairs included in the structure, a different one for all the remaining possible pairs.\\
\textbf{}   & It also draws the given consensus structure annotated with the significantly covarying base pairs.\\
 & \\ 
\end{tabular}

\subsection {The four options to run \rscape}

These are the four options to run R-scape.

\begin{labeling}{Evaluate region for conserved structure}
  
\item[Evaluate region for conserved structure]

  All possible pairs are analyzed equally as a one set test.  If a
  consensus structure is provided, that structure is ignored in the
  covariation test, but it is visualized with the significant
  covarying pairs highlighted in green.

  \textbf{preferred use:}\\
  This option is most appropriate if you're trying to determine if a conserved structure exists.\\
  
\item[Predict new structure]

  All possible pairs are analyzed equally.
  A structure is predicted and visualized with the significant covarying pairs highlighted in green.

  \textbf{preferred use:}\\
  This option is most appropriate for obtaining a new consensus
  structure prediction based on covariation analysis.

\item[Evaluate given structure]
  
  Requires that your Stockholm file has a proposed consensus structure annotation.
  Two independent covariation tests are performed, one on the set of proposed basepairs, the other on all other possible pairs.
  The given structure is visualized with significant covarying pairs highlighted in green.

  \textbf{preferred use:}\\
  This option is most appropriate for evaluating how well an
  independently proposed consensus structure is supported by
  covariation analysis.


\item[Improve given structure]

  Requires that your Stockholm file has a proposed consensus structure annotation.
  Two independent covariation tests are performed, one on the set of proposed basepairs, the other on all other possible pairs.
  A new consensus structure is predicted and visualized with the significant covarying pairs highlighted in green.

  \textbf{preferred use:}\\
  This option is most appropriate for using covariation analysis to improve your current consensus structure.

\end{labeling}



I'll show examples of running each mode, using examples in the
\ccode{tutorial/} subdirectory of the distribution.


\subsection {Option --RAF(S) disallowed}

The options to use the covariation measures RAF, RAFa, RAFp, RAFS, RAFSp, and RAFSa has
been disallowed, unless they are used in combination with option
--naive which reports the list of values for all possible pairs
without any statistical significance associated to them.

The following disclamer appears otherwise.

\user{bin/R-scape --RAF tutorial/updated\_Arisong.sto}\\

\begin{sreoutput}
DISCLAIMER: This measure can only be used in combination with the --naive option.

The --naive option reports a ranked list of scores for all possible
pairs without assigning E-values. RAF, RAFS and related measures
(RAFp, RAFa, RAFSp, RAFSa) cannot be used in combination with
R-scape's statistical test.

The RAF(S) statistics measure covariation and consistency. RAFS
assigns relatively high scores to pairs of alignment columns that are
consistent with base pairing even if there is no covariation at
all. The RAFS statistic was developed for the purpose of predicting
consensus RNA structures from alignments of sequences already presumed
to have a structure (Hofacker et al., 2002; Lindgreen et al.,
2006). For this purpose, both covariation and consistency are useful
cues. Distinguishing a conserved RNA structure from a conserved
primary sequence is a different problem that requires using a
statistic that does not systematically detect significant signals on
conserved primary sequence alone. That is R-scape's statistical
test. The R-scape statistical test can only be used with measures that
estimate covariation alone such as mutual information (MI) or G-test
(GT).
\end{sreoutput}

\subsection{Files used in the tutorial}

The subdirectory \prog{/tutorial} in the \rscape\ distribution contains the
files used in the tutorial. 

The tutorial provides several examples of RNA structural
alignments, all in Stockholm format:

\begin{sreitems}{\emprog{updated\_Arisong.sto}}
\item[\emprog{updated\_Arisong.sto}] Structural alignment of the ciliate
  Arisong RNA. This alignment is an updated
  version of the one published in~\citep{JungEddy11}.
\item[\emprog{ar14.sto}] Structural alignment of the $\alpha$-proteobacteria ncRNA ar14. This alignment is an updated version of the one
  published in~\citep{delVal12}.
\item[\emprog{manA.sto}] Alignment of the manA RNA motif~\citep{Weinberg09,Weinberg10} provided in the Zasha Weinberg database (ZWD)~\citep{ZWD18}.
\item[\emprog{RF00005.sto}] Rfam v12.0~\citep{Nawrocki15} seed alignment of tRNA. 
\item[\emprog{RF00001-noss.sto}] Rfam v12.0 seed alignment of 5S rRNA, after removing the consensus secondary structure. 
\end{sreitems}


\subsection{Running \rscape\, on one alignment file}
To run \rscape\ with default parameters on alignment file
\prog{tutorial/updated\_Arisong.sto} use:\\

\user{bin/R-scape tutorial/updated\_Arisong.sto}\\

\noindent
The output is a list of the significantly covarying positions under the one-set test

\begin{sreoutput}
# R-scape :: RNA Structural Covariation Above Phylogenetic Expectation
# R-scape 2.6.10 (July 2026)
# Copyright (C) 2017-2025 Elena Rivas, Harvard University.
# Freely distributed under the GNU General Public License (GPLv3).
#-------------------------------------------------------------------------------------------------------
# MSA updated_Arisong_1 nseq 95 (95) alen 66 (150) avgid 65.82 (64.97) nbpairs 20 (20)
# Two-set statistical test (one test for annotated basepairs, another for all other pairs)
#
#
# Method Target_E-val [cov_min,cov_max] [FP | TP True Found | Sen PPV F] 
# GTp    0.05         [-9.95,121.66]     [0 | 11 20 11 | 55.00 100.00 70.97] 
#
#  left_pos    right_pos   score       E-value        pvalue       substitutions    power   nseff(%)        % canonical_pairs
#-------------------------------------------------------------------------------------------------------------------------------------------------
*        98          106   121.65824   9.48364e-12    4.74182e-13   8               0.28    76.48 (80.5%)   100.0
*       122          137   91.44681    1.53068e-07    7.65341e-09   22              0.66    87.92 (92.5%)   99.2
\end{sreoutput}
A star ``*'' in the first column indicates that the pair is part of
the annotated structure in the \prog{updated\_Arisong.sto} file. A
blank indicates a pair that is not compatible with the structure. A
``$\sim$`` indicates an interaction not in the annotated structure but
compatible with it (none in this example).

The Arisong RNA in \prog{tutorial/updated\_Arisong.sto} has a proposed
secondary structure.  Instead of testing all pairs as equivalent, we
may want to test the significance of the given structure as a one set
of pairs, and independently that of the rest of all possible pairs.
In order to do a two-set test use:\\

\user{bin/R-scape -s tutorial/updated\_Arisong.sto}\\

\noindent
The output is a list of the significantly covarying positions under the two-set test.

\begin{sreoutput}
# R-scape :: RNA Structural Covariation Above Phylogenetic Expectation
# R-scape 2.6.10 (July 2026)
# Copyright (C) 2017-2025 Elena Rivas, Harvard University.
# Freely distributed under the GNU General Public License (GPLv3).
#-------------------------------------------------------------------------------------------------------
# MSA updated_Arisong_1 nseq 95 (95) alen 66 (150) avgid 65.82 (64.97) nbpairs 20 (20)
# Two-set statistical test (one test for annotated basepairs, another for all other pairs)
#
#
# Method Target_E-val [cov_min,cov_max] [FP | TP True Found | Sen PPV F] 
# GTp    0.05         [-9.95,121.66]     [0 | 11 20 11 | 55.00 100.00 70.97] 
#
#  left_pos   right_pos   score       E-value      pvalue       substitutions   power   nseff(%)       % canonical_pairs
#--------------------------------------------------------------------------------------------------------------------------------------------------
*        98        106    121.65824   9.48364e-12  4.74182e-13     8            0.28    76.48 (80.5%)   100.0
*       122        137     91.44681   1.53068e-07  7.65341e-09     22           0.66    87.92 (92.5%)   99.2
*        96        108     88.43485   3.96567e-07  1.98284e-08     6            0.19    84.22 (88.6%)   100.0
*       120        139     74.80346   2.84591e-05  1.42296e-06     33           0.79    93.46 (98.4%)   92.7
*       119        140     58.72199   0.00348352   0.000174176     31           0.77    80.36 (84.6%)   90.8
*       121        138     58.34868   0.00388111   0.000194056     27           0.73    86.53 (91.1%)   89.7
*        94        110     57.28027   0.00511565   0.000255783     12           0.42    87.61 (92.2%)   97.3
*       124        134     55.67759   0.00754897   0.000377449     13           0.45    73.10 (76.9%)   100.0
*       123        135     54.59659   0.00943824   0.000471912     20           0.62    81.32 (85.6%)   100.0
*        99        105     53.44824   0.0111888    0.000559441     6            0.19    88.96 (93.6%)   99.3
*        97        107     44.91860   0.044289     0.00221445      6            0.19    81.87 (86.2%)   100.0


# Powertype = SINGLE SUBSTITUTIONS
# covary  left_pos      right_pos    substitutions      power   nseff(%)       % canonical_pairs
#-------------------------------------------------------------------------------------------------
     *    94            110             12              0.42    87.61 (92.2%)   97.3
          95            109             9               0.32    77.90 (82.0%)   97.3
     *    96            108             6               0.19    84.22 (88.6%)   100.0
     *    97            107             6               0.19    81.87 (86.2%)   100.0
     *    98            106             8               0.28    76.48 (80.5%)   100.0
     *    99            105             6               0.19    88.96 (93.6%)   99.3
          100           104             2               0.00    61.10 (64.3%)   100.0
          111           148             0               0.00    79.55 (83.7%)   100.0
          112           147             6               0.19    79.98 (84.2%)   100.0
          113           146             1               0.00    74.68 (78.6%)   100.0
          114           145             3               0.04    84.37 (88.8%)   100.0
          115           144             14              0.48    86.25 (90.8%)   98.9
          116           143             37              0.82    90.29 (95.0%)   60.9
     *    119           140             31              0.77    80.36 (84.6%)   90.8
     *    120           139             33              0.79    93.46 (98.4%)   92.7
     *    121           138             27              0.73    86.53 (91.1%)   89.7
     *    122           137             22              0.66    87.92 (92.5%)   99.2
     *    123           135             20              0.62    81.32 (85.6%)   100.0
     *    124           134             13              0.45    73.10 (76.9%)   100.0
          125           133             12              0.42    81.58 (85.9%)   91.7
#
# BPAIRS 20
# avg substitutions per BP  13.4
# BPAIRS expected to covary 7.6 +/- 1.8
# BPAIRS observed to covary 11

#
# RMs = 3 L = 150

# RM_HELIX NESTED 94-100 104-110, nbp = 7 nbp_cov = 5
# aggregated NONE
n/a
NESTED  :::::::::::<<<<<<<___>>>>>>>::::::::::::::::::::::::::::::::::::::
        :::::::::::<-<<<<_____>>>>->::::::::::::::::::::::::::::::::::::::
NESTED  :::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::<<<<<<<___>>>>>>>::::::::::::::::::::::::::::::::::::::::
        :::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::<-<<<<_____>>>>->::::::::::::::::::::::::::::::::::::::::

# RM_HELIX NESTED 119-125 133-140, nbp = 7 nbp_cov = 6
# aggregated NONE
n/a
NESTED  ::::::::::::::::::::::::::::::::::::<<<<<<<_____>>>->>>>::::::::::
        ::::::::::::::::::::::::::::::::::::<<<<<<_______>>->>>>::::::::::
NESTED  ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::<<<<<<<_______>>>->>>>::::::::::
        ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::<<<<<<_________>>->>>>::::::::::

# RM_HELIX NESTED 111-116 143-148, nbp = 6 nbp_cov = 0
# aggregated NONE
n/a
NESTED  ::::::::::::::::::::::::::::<<<<<<________________________>>>>>>::
        ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
NESTED  ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::<<<<<<__________________________>>>>>>::
        ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# Number of covarying pairs = 11

# MSA updated_Arisong_1 alen=66 nseq=95 time=0.06 secs
\end{sreoutput}
The scores of the pairs are identical to those in the one-set
test. The E-values have changed relative to those of the one-set test.


 \subsection{Default parameters}

Default parameters are:

\begin{sreitems}{\emprog{Pairwise percent identity:}}
\item[\emprog{Target E-value:}]default is 0.05. \rscape\, reports
  pairs which covariation score has E-value smaller or equal to the
  target value.  The target E-value can be changed with option
  \emprog{-E <x>}, $x >= 0$.

\item[\emprog{Sequence weighting:}]Sequences are weighted according to
  the Gerstein/Sonnhammer/Chothia (GSC)
  algorithm~\citep{Gerstein94}. This algorithm is time consuming. For
  alignments with more than 1000 sequences, we use the faster
  position-based weighting algorithm~\citep{Henikoff94b}. Both
  weighting algorithms are implemented as part of the easel library.

\item[\emprog{Gaps in columns:}]Columns with more than 50\% gaps are
  removed. The gap threshold for removing columns can be modified
   using option \emprog{--gapthresh <x>} , $0<x<=1$.

 \item[\emprog{Covariation statistic:}]The default covariation statistic
   is the average product corrected G-Test (equivalent to option
   \emprog{--GTp}).

 \item[\emprog{Covariation Class:}]\rscape\ uses the 16 component
   covariation statistic (C16), unless the number of sequences in the
   alignment is $\leq$ 8 or the length of the alignment is $\leq$ 50,
   in which case it uses the two-class covariation statistic (C2). A
   particular covariation class can be selected using either
   \emprog{--C16} or \emprog{--C2}.

   The threshold for the minimum number of sequences can be changed
   with option \prog{--nseqthresh <n>}.  The threshold for the minimum
   alignment length can be changed with option \prog{--alenthresh <n>}.

 \item[\emprog{Null alignments:}]In order to estimate E-values,
   \rscape\ produces 20 null alignments, unless the product of the
   number of sequences by the length of the alignment $<$ 10,000 in
   which case the number of null alignments is 50; or $<$ 1,000 in
   which case it is 100. The number of null alignments can be
   controlled with option \emprog{--nshuffle <n>}.
 \end{sreitems}

 A full list of the \rscape\ options is found by using

 \user{\rscape\ -h}

 
 
 
